---
title: "Shiny App Submission Program Preparation"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
    toc_float: true
resource_files:
   - pdf/*.pdf
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(pkglite)
library(purrr)
```


```{r, message = FALSE}
# Initiate start-up file
source(file.path(rprojroot::find_root("DESCRIPTION"), "inst/startup.R"))
```
# Introduction

This document summarizes the preparation and installation instructions for the R Consortium Submission Pilot 2 Shiny application. Here is a brief description of the procedures:

* The application files are assembled into a customized text file using the [`{pkglite}`](https://merck.github.io/pkglite/) R package, with the intent of using this text file to submit the application through the FDA's submission portal based on the FDA's [Electronic Common Technical Document](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd) (eCTD) format.
* Deploy the application locally by unpacking the app source code from the aforementioned text file. Since the Shiny application is created as an R package using the [`{golem}`](https://thinkr-open.github.io/golem/) package, the application can be easily run with a single function call.

## Application Dependencies and Environment

The Shiny application in this pilot was created within R version 4.2.0, and the [`{renv}`](https://rstudio.github.io/renv/index.html) package was used to manage the application dependencies to maintain a reproducible environment. All packages and their associated versions are listed in the table below.

```{r, eval = TRUE}
pkg <- desc::desc_get_deps() %>%
  filter(type == "Imports") %>%
  pull(package)

# custom code to obtain package title and version
pkg_desc <- tibble::tibble(Package = pkg) %>%
  mutate(Title = purrr::map_chr(Package, ~utils::packageDescription(.x, fields = "Title"))) %>%
  mutate(Version = purrr::map_chr(Package, ~utils::packageDescription(.x, fields = "Version"))) %>%
  arrange(Package)

pkg_desc
```

## Pack Shiny Application

The code below assembles the application's code and supporting files into a custom text file using `{pkglite}` and saved into `ectd/r1pk1.txt`. Please note that the size of this text file is approximately 37 megabytes due to the included data sets within the application.

```{r}
# specification for app.R file
app_spec <- file_spec(
  ".", 
  pattern = "\\.R$", 
  format = "text", 
  recursive = FALSE
)

app_source_spec <- file_spec(
  "R",
  pattern = "\\.R",
  format = "text",
  recursive = FALSE
)

renv_spec <- file_spec(
  ".",
  pattern = "\\.lock",
  format = "text",
  recursive = FALSE
)

renv_spec2 <- file_spec(
  "renv",
  pattern = "\\.R$",
  format = "text",
  recursive = FALSE
)

renv_spec3 <- file_spec(
  "renv/cellar",
  pattern = "\\.tar.gz|\\.zip",
  format = "binary",
  recursive = TRUE
)

golem_spec <- file_auto("dev")

file_ectd2 <- file_spec(
  ".",
  pattern = "^DESCRIPTION$|^NAMESPACE$|^README$|^README\\.md$|^NEWS$|^NEWS\\.md$|^LICENSE$|\\.Rbuildignore$|\\.Rprofile$|\\.Renviron$",
  format = "text",
  recursive = FALSE,
  ignore_case = FALSE,
  all_files = TRUE
)

path$home %>%
  collate(
    file_ectd2,
    file_auto("inst"), 
    app_spec, 
    app_source_spec,
    renv_spec,
    renv_spec2,
    renv_spec3,
    golem_spec
  ) %>%
  pack(output = "ectd/r1pkg.txt")
```

## Application Installation (Dry run) 

To unpack the application source code and supporting files from the custom `ectd/r1pkg.txt` bundle and run the Shiny application, perform the following steps in your preferred R development environment, such as RStudio or a R console session:

1. Ensure that the working directory is set to the same location as the downloaded `r1pkg.txt` file, and run the following code: `pkglite::unpack(input = "r0pkg.txt", install = FALSE)`. You should see a series of messages indicating the application files being written to a subdirectory called `pilot2wrappers`.
1. Launch a new R session in the `pilot2wrappers` subdirectory. 
1. Ensure that the package library is restored with `{renv}` by running the following code: `renv::restore()`. After confirmation, the packages used by the application will be downloaded and installed in a project-specific library.
1. To run the application, you have the following choices. If you are using RStudio to view the application, open the `app.R` file and click the Run App button at the top of the file window. If you are using an R console, run the following code: `golem:::run_dev()`. 
